---
title: 'Too Many Facets?: A Tidy Pipeline to a slickR Slideshow'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

It's almost banal to say that facets in ggplot2 are an immensely useful tool. `facet_wrap` and `facet_grid` changed how I approach analysis and visualization, but I occasionally run into the issue of having too many facets.

- There may be so many subgroups in the data that not even playing with `fig.height` can fit everything in.
- Each graph is relatively detailed and faceting distorts how they look.
- When exploring, I want to view each variable or subgroup on its own.

When working on a recent project where I wanted to display a separate graph for each US state, I personally discovered [slickR](https://github.com/metrumresearchgroup/slickR), an R interface to the [slick](http://kenwheeler.github.io/slick/) JavaScript carousel library. Combining this with the tidyverse tool chain, it was (relatively) straightforward to go from  data to a slideshow.

The code below builds off the canonical gapminder example from the [tidyr 0.4.0 release blog post](https://blog.rstudio.com/2016/02/02/tidyr-0-4-0/) and the [slickR](https://metrumresearchgroup.github.io/slickR) examples to create a slideshow with a separate graph for each country in the data set.

```{r}
# Data
library(gapminder)

# Tidy toolchain
library(dplyr)
library(tidyr)
library(purrr)

# Visualize
library(ggplot2)

# For the slideshow 
library(svglite)
library(slickR)
```

As a reminder, gapminder includes the life expectancy, GDP per capita, and population of countries over the last half-century.

```{r}
gapminder
```

In this example, I will graph the life expectancy over time for each country. The first section of code follows the classic `group_by() %>% nest() %>% mutate(map())` approach. The second section updates the slickR example to use purrr. 

While it's possible to pipe directly into `slickR`, I prefer saving the vector of graphs. This allows for adjusting the settings on slickR (when working interactively) or saving the object for use elsewhere. (The graphs are created in a script and used in a dashboard or report.)

```{r country_graphs, cache = TRUE}
gapminder %>% 
  # Use the group_by %>% nest pattern to group data by country
  group_by(continent, country) %>% 
  nest() %>% 
  # Use map2 so the country name can be used as the title
  mutate(graphs = map2(data, country,
                       ~ggplot(data = .x, aes(year, lifeExp)) +
                         geom_point() +
                         geom_line() +
                         labs(title = .y)
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) %>% 
  # Convert to character vector
  map_chr(function(sv) paste0('data:image/svg+xml;utf8,', as.character(sv))) ->
  country_graphs

```

```{r}
slickR(country_graphs, slickOpts = list(dots = FALSE))
```

By default, `slickR` displays one graph per slide and includes a dot for each. There are a number of other options within `slickOpts`, including the number of objects to include in each slide.

```{r}
sessionInfo()
```

